---
title: "Non-disclosive federated exposome data analysis with DataSHIELD and Bioconductor"
author:
- name: Xavier Escribà Montagut
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  email: xavier.escriba@isglobal.org
- name: Juan R. Gonzalez
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  email: juanr.gonzalez@isglobal.org
date: "`r Sys.Date()`"
package: "`r pkg_ver('dsExposome')`, `r pkg_ver('dsExposomeClient')`"
output:
  BiocStyle::html_document:
    number_sections: true
    toc: yes
    fig_caption: yes
    toc_float: true
bibliography: dsExposome_Vignette.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment="", warning=FALSE, message=FALSE, cache=FALSE)
```

# Purpose

# Setup

# Exposome analysis

## Loading the Exposome Set

An exposome dataset is made up from [three different tables](https://www.bioconductor.org/packages/release/bioc/vignettes/rexposome/inst/doc/exposome_data_analysis.html#data-format). In order to be studied, this three tables need to be coerced into a R object of class `ExposomeSet`, for that reason an exposome dataset can be available on a server as three tables or as a resource.

### From tables to `ExposomeSet`

To coerce the three tables that make the exposome dataset into an `ExposomeSet` R object, the tables have to be loaded inside the same study server.

```{r}
library(DSLite)
library(dsBaseClient)
library(dsExposome)

exposures <- utils::read.table("../test_data/exposures.csv", header = TRUE,
                         row.names = "idnum", sep = ",", na.strings = c("NA", "-", "?", " ", ""))
phenotypes <- utils::read.table("../test_data/phenotypes.csv", header = TRUE,
                         row.names = "idnum", sep = ",", na.strings = c("NA", "-", "?", " ", ""),
                         stringsAsFactors = TRUE)
description <- utils::read.table("../test_data/description.csv", header = TRUE,
                          row.names = "Exposure", sep = ",", na.strings = c("NA", "-", "?", " ", ""))

dslite.server <- newDSLiteServer(tables=list(exposures = exposures, description = description, phenotypes = phenotypes),
                                 config = DSLite::defaultDSConfiguration(include=c("dsBase", "dsExposome")))

log <- data.frame(server = c("sim1"),
                  url = c("dslite.server"),
                  table = list("exposures", "description", "phenotypes"),
                  driver = c("DSLiteDriver"))
conns <- datashield.login(log, assign = T)

ds.ls()
```

To coerce the three tables, the `dsExposomeClient` library has the `ds.loadExposome()` function, which takes as input parameters the names of the tables on the study server among some further configuration options, refer to the function documentation for more information.

```{r}
library(dsExposomeClient)

ds.loadExposome("exposures", "description", "phenotypes", "Family", 5, FALSE, "exposome_set")
ds.class("exposome_set")

datashield.logout(conns)
```

### From resource to `ExposomeSet`

When there's an `ExposomeSet` resource available, it just needs to be loaded into the study server

```{r}
library(DSLite)
library(dsBaseClient)
library(dsExposome)
library(resourcer)
library(rexposome)

exposome <- newResourceClient(
  resourcer::newResource(
    name = "exposome",
    url = "https://github.com/isglobal-brge/dsExposomeClient/raw/master/test_data/exposome_resource.Rdata",
    format = "ExposomeSet"
  )
) 

dslite.server <- newDSLiteServer(resources=list(exposome),
                                 config = DSLite::defaultDSConfiguration(include=c("dsBase", "dsExposome", "resourcer", "rexposome")))

log <- DSI::newDSLoginBuilder()
log$append(server = "sim1", url = "dslite.server", resource = "exposome", driver = "DSLiteDriver")
log <- log$build()

conns <- datashield.login(log, assign = T)

ds.ls()
ds.class("exposome")
```

Once on the study server, the resource has to be coerced into an R object

```{r}
datashield.assign.expr(conns, symbol = "exposome_object",
                       expr = quote(as.resource.object(exposome)))
ds.class("exposome_object")
```

## Performing an ExWAS

When the `ExposomeSet` object is on the study the server, the function `ds.exwas` is in charge of performing the ExWAS. As an example let's present a situation where each exposition has to be associated to the `blood_pre` phenotype using the phenotype `sex` as a covariate. This study would be done as follows

```{r}
library(dsExposomeClient)

exwas_results <- ds.exwas("blood_pre ~ sex", "exposome_object", "gaussian")
head(exwas_results)
```

The model is inputed as a string, where the left side term is the objective phenotype, and the right term are the covariates, in the case of wanting more covariates proceed as: `objective ~ cov1 + cov2 + ... + covN`. It's important noting that if the output family (third argument) does not match the objective family, the ExWAS will fail (Example: The objective family is `binomial` and the user sets it to `gaussian`).
