---
title: "Non-disclosive federated exposome data analysis with DataSHIELD and Bioconductor"
author:
- name: Xavier Escribà Montagut
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  email: xavier.escriba@isglobal.org
- name: Juan R. Gonzalez
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  email: juanr.gonzalez@isglobal.org
date: "`r Sys.Date()`"
package: "`r pkg_ver('dsExposome')`, `r pkg_ver('dsExposomeClient')`"
output:
  BiocStyle::html_document:
    number_sections: true
    toc: yes
    fig_caption: yes
    toc_float: true
bibliography: dsExposome_Vignette.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment="", warning=FALSE, message=FALSE, cache=FALSE)
```

# Purpose

# Setup

# Exposome analysis

## Loading the Exposome Set

An exposome dataset is made up from [three different tables](https://www.bioconductor.org/packages/release/bioc/vignettes/rexposome/inst/doc/exposome_data_analysis.html#data-format). In order to be studied, this three tables need to be coerced into a R object of class `ExposomeSet`, for that reason an exposome dataset can be available on a server as three tables or as a resource.

### From tables to `ExposomeSet`

To coerce the three tables that make the exposome dataset into an `ExposomeSet` R object, the tables have to be loaded inside the same study server.

```{r}
library(DSLite)
library(dsBaseClient)
library(dsExposome)

exposures <- utils::read.table("../test_data/exposures.csv", header = TRUE,
                         row.names = "idnum", sep = ",", na.strings = c("NA", "-", "?", " ", ""))
phenotypes <- utils::read.table("../test_data/phenotypes.csv", header = TRUE,
                         row.names = "idnum", sep = ",", na.strings = c("NA", "-", "?", " ", ""),
                         stringsAsFactors = TRUE)
description <- utils::read.table("../test_data/description.csv", header = TRUE,
                          row.names = "Exposure", sep = ",", na.strings = c("NA", "-", "?", " ", ""))

dslite.server <- newDSLiteServer(tables=list(exposures = exposures, description = description, phenotypes = phenotypes),
                                 config = DSLite::defaultDSConfiguration(include=c("dsBase", "dsExposome")))

log <- data.frame(server = c("sim1"),
                  url = c("dslite.server"),
                  table = list("exposures", "description", "phenotypes"),
                  driver = c("DSLiteDriver"))
conns <- datashield.login(log, assign = T)

ds.ls()
```

To coerce the three tables, the `dsExposomeClient` library has the `ds.loadExposome()` function, which takes as input parameters the names of the tables on the study server among some further configuration options, refer to the function documentation for more information.

```{r}
library(dsExposomeClient)

ds.loadExposome("exposures", "description", "phenotypes", "Family", 5, FALSE, "exposome_set")
ds.class("exposome_set")

datashield.logout(conns)
```

### From resource to `ExposomeSet`

When there's an `ExposomeSet` resource available, it just needs to be loaded into the study server

```{r}
library(DSLite)
library(dsBaseClient)
library(dsExposome)
library(resourcer)
library(rexposome)

exposome <- newResourceClient(
  resourcer::newResource(
    name = "exposome",
    url = "https://github.com/isglobal-brge/dsExposomeClient/raw/master/test_data/exposome_resource.Rdata",
    format = "ExposomeSet"
  )
) 

dslite.server <- newDSLiteServer(resources=list(exposome),
                                 config = DSLite::defaultDSConfiguration(include=c("dsBase", "dsExposome", "resourcer", "rexposome")))

log <- DSI::newDSLoginBuilder()
log$append(server = "sim1", url = "dslite.server", resource = "exposome", driver = "DSLiteDriver")
log <- log$build()

conns <- datashield.login(log, assign = T)

ds.ls()
ds.class("exposome")
```

Once on the study server, the resource has to be coerced into an R object

```{r}
datashield.assign.expr(conns, symbol = "exposome_object",
                       expr = quote(as.resource.object(exposome)))
ds.class("exposome_object")
```

## Descriptive statistics of the exposome dataset

Non-disclosive descriptive statistics can be obtained from the exposome dataset, the function `ds.exposome_summary` is in charge of that. It can obtain descriptive statistics from numeric and factor variables of the exposome dataset (both from exposures and phenotypes). To obtain all the exposure and phenotypes of the exposome set

```{r}
ds.exposome_variables("exposome_object" , "exposures", conns)
ds.exposome_variables("exposome_object" , "phenotypes", conns)
```

To get the descriptive statistics of the variable of interest 

```{r}
# Numerical variable
ds.exposome_summary("exposome_object", "PM25", conns)
# Factor variable
ds.exposome_summary("exposome_object", "flu", conns)
```

## Performing an ExWAS

When the `ExposomeSet` object is on the study the server, the function `ds.exwas` is in charge of performing the ExWAS. As an example let's present a situation where each exposition has to be associated to the `blood_pre` phenotype using the phenotype `sex` as a covariate. This study would be done as follows

```{r}
library(dsExposomeClient)

exwas_results <- ds.exwas("blood_pre ~ sex", "exposome_object", "gaussian", FALSE, conns)
head(exwas_results$exwas_results)
```

The model is inputed as a string, where the left side term is the objective phenotype, and the right term are the covariates, in the case of wanting more covariates proceed as: `objective ~ cov1 + cov2 + ... + covN`. It's important noting that if the output family (third argument) does not match the objective family, the ExWAS will fail (Example: The objective family is `binomial` and the user sets it to `gaussian`).

## Exposures PCA

A PCA can be performed on the exposures of the exposome dataset. To do so, there's the `ds.exposome_pca`. The function offers the functionality of standarizing the input data with the argument `standar`. It is important noting that this function is sensitive to be disclosive, specially for very rectangular data frames (similar number of variables as individuals). To illustrate this problem, let's try to perform a PCA on the whole exposures test data.

```{r error = TRUE}
ds.exposome_pca("exposome_object", TRUE)
```

If that is the case, one option is to reduce the families of exposures of the Exposome Set. There's an already subsetted Exposome Set object available on the `test_data` folder of `dsExposomeClient` for demostration purposes.

```{r}
exposome_pca <- newResourceClient(
  resourcer::newResource(
    name = "exposome",
    url = "https://github.com/isglobal-brge/dsExposomeClient/raw/master/test_data/exposome_resource_pca.Rdata",
    format = "ExposomeSet"
  )
) 

dslite.server <- newDSLiteServer(resources=list(exposome_pca),
                                 config = DSLite::defaultDSConfiguration(include=c("dsBase", "dsExposome", "resourcer", "rexposome")))

log <- DSI::newDSLoginBuilder()
log$append(server = "sim1", url = "dslite.server", resource = "exposome_pca", driver = "DSLiteDriver")
log <- log$build()

conns <- datashield.login(log, assign = T)

datashield.assign.expr(conns, symbol = "exposome_object_pca",
                       expr = quote(as.resource.object(exposome_pca)))

pca <- ds.exposome_pca("exposome_object_pca")
head(pca[,1:4])
```

