---
title: "Non-disclosive federated exposome data analysis with DataSHIELD and Bioconductor"
author:
- name: Xavier Escribà Montagut
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  email: xavier.escriba@isglobal.org
- name: Juan R. Gonzalez
  affiliation:
  - &isglobal Bioinformatics Research Group in Epidemiolgy (BRGE), Barcelona Insitute for Global Health (ISGlobal)
  email: juanr.gonzalez@isglobal.org
date: "`r Sys.Date()`"
package: "`r pkg_ver('dsExposome')`, `r pkg_ver('dsExposomeClient')`"
output:
  BiocStyle::html_document
bibliography: dsExposome_Vignette.bib
vignette: >
  %\VignetteIndexEntry{Non-disclosive federated exposome data analysis with DataSHIELD and Bioconductor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
BiocStyle::markdown()
options(width=80)
knitr::opts_chunk$set(comment="", warning=FALSE, message=FALSE, cache=FALSE)
```

# Purpose

# Setup

# Exposome analysis

## Connecting to the Opal server



## Loading the Exposome Set

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/loading_data.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/loading_data.Rmd"))
```

```{r child='/chapter1/loading_data.Rmd', include=FALSE}
```

### From tables to `ExposomeSet`

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/tables_to_exposome.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/tables_to_exposome.Rmd"))
```

```{r child='/chapter1/tables_to_exposome.Rmd', include=FALSE}
```

### From resource to `ExposomeSet`

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/resource_to_exposome.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/resource_to_exposome.Rmd"))
```

```{r child='/chapter1/resource_to_exposome.Rmd', include=FALSE}
```

## Exploring the loaded dataset

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/exploring_dataset.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/exploring_dataset.Rmd"))
```

```{r child='/chapter1/exploring_dataset.Rmd', include=FALSE}
```

### Family names

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/family_names.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/family_names.Rmd"))
```

```{r child='/chapter1/family_names.Rmd', include=FALSE}
```

### Exposures and phenotypes names

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/exp_pheno_names.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/exp_pheno_names.Rmd"))
```

```{r child='/chapter1/exp_pheno_names.Rmd', include=FALSE}
```

### Summary of variables

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/summary_variables.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/summary_variables.Rmd"))
```

```{r child='/chapter1/summary_variables.Rmd', include=FALSE}
```

### Missing data

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/missing_data.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/missing_data.Rmd"))
```

```{r child='/chapter1/missing_data.Rmd', include=FALSE}
```

### Exposures Normality

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/exposures_normality.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/exposures_normality.Rmd"))
```

```{r child='/chapter1/exposures_normality.Rmd', include=FALSE}
```

### Exposures Behaviour

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/exposures_behaviour.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/exposures_behaviour.Rmd"))
```

```{r child='/chapter1/exposures_behaviour.Rmd', include=FALSE}
```

## Exposures Imputation

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/exposures_imputation.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/exposures_imputation.Rmd"))
```

```{r child='/chapter1/exposures_imputation.Rmd', include=FALSE}
```

## Performing an ExWAS

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/exwas.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/exwas.Rmd"))
```

```{r child='/chapter1/exwas.Rmd', include=FALSE}
```

## Exposures PCA

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/pca.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/pca.Rmd"))
```

```{r child='/chapter1/pca.Rmd', include=FALSE}
```

## Exposures Correlation

```{r include = FALSE}
download.file("https://github.com/isglobal-brge/dsExposomeClient/raw/master/vignette/correlation.Rmd", paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/chapter1/correlation.Rmd"))
```

```{r child='/chapter1/correlation.Rmd', include=FALSE}
```