The number of missing data on each exposure and on each phenotype can be found by using the function `ds.tableMissings`. This function returns a vector with the amount of missing data in each exposure or phenotype. The argument `set` indicates if the number of missing values is counted on exposures or phenotypes. The argument `output` indicates if it is shown as counts (`output="n"`) or as percentage (`output="p"`).

```{r}
ds.tableMissings("exposome_object", set = "exposures")
ds.tableMissings("exposome_object", set = "phenotypes")
```

Optionally, there's also the `ds.plotMissings` function which returns a `ggplot` object with a barplot of missings for exposures or phenotypes, there's the option of displaying the percentage of missings as well as total counts, check the function documentation for further information.

```{r}
ds.plotMissings("exposome_object", "exposures", "p")
```

Please note that since this function call returns a `ggplot` object, it can be plotted properly to avoid squished Y axis when there are lots of exposures, read the [official documentation](https://ggplot2.tidyverse.org/reference/ggsave.html) for information on how to do that.