Most of the test done in *exposome* analysis requires that the exposures must follow a normal distribution. The function `ds.normalityTest` performs a test on each exposure for normality behaviour. The result is a `data.frame` with the exposures’ names, a flag `TRUE`/`FALSE` for normality and the p-value obtained from the *Shapiro-Wilk* Normality Test (if the p-value is under the threshold, then the exposure is not normal).

```{r}
nm <- ds.normalityTest("exposome_object")
table(nm$server1$normality)
```

So, the exposures that do not follow a normal distribution are:
  
```{r}
nm$server1$exposure[!nm$server1$normality]
```

The `ds.normalityTest` function has some extra input arguments to tune the normality test, check the function documentation for further information.

The exposures can be visualized using non-disclosive histograms to see their distribution along their *Shapiro-Wilk* Normality Test p-value.

```{r results = FALSE}
ds.exposure_histogram("exposome_object", "AbsPM25")
```

If the selected exposure is not following a normal distribution, the function `ds.exposure_histogram` accepts the argument `show.trans` to visualize the raw data histogram plus three typical transformations (`exp`, `log` and `sqrt`), the *Shapiro-Wilk* Normality Test p-value is shown for all the transformations.

```{r results = FALSE}
ds.exposure_histogram("exposome_object", "AbsPM25", TRUE)
```