The correlation between exposures, in terms of intra-family and inter-family exposures, is interesting to take into account. The correlation of the exposome can be computed using `ds.correlation`. The correlation could be disclosive, for that reason there's the argument `fam` to select the families to compute the correlation subsetting the exposome dataset by families, typical complete exposome datasets will fail as the correlation matrix could be disclosive.

```{r}
ds.correlation("exposome_object")[[1]][[1]]$`Correlation Matrix`[1:5,1:5]
ds.correlation("exposome_object")[[1]][[1]]$`Error message`
```

When subsetting the exposome set by families the correlation succeeds.

```{r}
ds.correlation("exposome_object", c("Metals", "Noise"))[[1]][[1]]$`Correlation Matrix`[1:5,1:5]
```

The output from the `ds.correlation` consists of the correlation matrix and a vector that contains the names of exposures and families on the correlation table, this second item is used by the `corPlot` function to display the results of the correlation study.

```{r}
corr_data <- ds.correlation("exposome_object", c("Metals", "Noise"))
corPlot(corr_data)
```

The best option to see the inter-family correlations is the *circos of correlations* while the *matrix of correlations* is a better way for studying the intra-family correlations.

```{r}
corPlot(corr_data, type = "circos")
```


