When the `ExposomeSet` object is on the study the server, the function `ds.exwas` is in charge of performing the ExWAS. As an example let's present a situation where each exposition has to be associated to the `blood_pre` phenotype using the phenotype `sex` as a covariate. This study would be done as follows

```{r}
exwas_results <- ds.exwas("blood_pre ~ sex", "exposome_object", "gaussian")
head(exwas_results$exwas_results)
```

This previous formula will fit linear models between the exposures and phenotype as follows:

```
    phenotype ~ exposure_1 + covar1 + ... + covarN
    phenotype ~ exposure_2 + covar1 + ... + covarN
    phenotype ~ exposure_3 + covar1 + ... + covarN
    ...
    phenotype ~ exposure_M + covar1 + ... + covarN
```

The model is inputted as a string, where the left side term is the objective phenotype, and the right term are the covariates, in the case of wanting more covariates proceed as: `objective ~ cov1 + cov2 + ... + covN`. It's important noting that if the output family (third argument) does not match the objective family, the ExWAS will fail (Example: The objective family is `binomial` and the user sets it to `gaussian`).

To visualize the results from the ExWAS, the function `ds.plotExwas` takes the output of `ds.exwas` and creates two different visualizations. A Manhattan-like plot with the p-values of the association between each exposure and asthma, coloured by families of exposures:
  
```{r fig.width=10, fig.height=12}
ds.plotExwas(exwas_results, "manhattan")
```

Also, a plot for the effects of a given model can be obtained with:
  
```{r fig.width=10, fig.height=12}
ds.plotExwas(exwas_results, "effect")
```

The ExWAS can also be applied as an inverse ExWAS, meaning that the linear models are the following:

```
    exposure_1 ~ phenotype + covar1 + ... + covarN
    exposure_2 ~ phenotype + covar1 + ... + covarN
    exposure_3 ~ phenotype + covar1 + ... + covarN
    ...
    exposure_M ~ phenotype + covar1 + ... + covarN
```

The funcion to perform this analysis is `ds.invExWAS`. The same example of the ExWAS would be written as:

```{r}
inverse_exwas_results <- ds.invExWAS(" ~ blood_pre + sex", "exposome_object", "gaussian")
head(inverse_exwas_results$exwas_results)
```

Note that in this case only right side term need to be supplied on the formula.

The same plotting functions can be used with the object returned by the inverse ExWAS analysis.

```{r}
ds.plotExwas(inverse_exwas_results, "manhattan")
```

