When the `ExposomeSet` object is on the study the server, the function `ds.exwas` is in charge of performing the ExWAS. As an example let's present a situation where each exposition has to be associated to the `blood_pre` phenotype using the phenotype `sex` as a covariate. This study would be done as follows

```{r}
exwas_results <- ds.exwas("blood_pre ~ sex", "exposome_object", "gaussian")
head(exwas_results$exwas_results)
```

The model is inputted as a string, where the left side term is the objective phenotype, and the right term are the covariates, in the case of wanting more covariates proceed as: `objective ~ cov1 + cov2 + ... + covN`. It's important noting that if the output family (third argument) does not match the objective family, the ExWAS will fail (Example: The objective family is `binomial` and the user sets it to `gaussian`).

To visualize the results from the ExWAS, the function `ds.plotExwas` takes the output of `ds.exwas` and creates two different visualizations. A Manhattan-like plot with the p-values of the association between each exposure and asthma, coloured by families of exposures:
  
  ```{r fig.width=10, fig.height=12}
ds.plotExwas(exwas_results, "manhattan")
```

Also, a plot for the effects of a given model can be obtained with:
  
  ```{r fig.width=10, fig.height=12}
ds.plotExwas(exwas_results, "effect")
```