The user can also prepare an `ExposomeSet` outside DataSHILED and provide this information in the Opal server as a *resource*. Our Opal demo (EXPOSOME project) has an `ExposomeSet` resource (see Figure \@ref(fig:resourcesExpo)) that can be loaded into the study server as follows:

```{r}
datashield.assign.resource(conns, symbol = 'exposome_resource', resource = list(server1 = 'EXPOSOME.exposomeSet'))

ds.ls()
ds.class("exposome_resource")
```

Once on the study server, the resource has to be coerced into an R object

```{r}
datashield.assign.expr(conns, symbol = "exposome_resource", expr = quote(as.resource.object(exposome_resource)))
ds.class("exposome_resource")
```

In that case, both `exposome_object` and `exposome_resource` are containing the same information.


```{r}
ds.exposome_variables("exposome_object", "phenotypes")
ds.exposome_variables("exposome_resource", "phenotypes")
```


```{r}
ds.exposome_summary("exposome_object", "AbsPM25")
ds.exposome_summary("exposome_resource", "AbsPM25")
```

```{r}
ds.exposome_summary("exposome_object", "blood_pre")
ds.exposome_summary("exposome_resource", "blood_pre")
```