We first start by loading both exposome and phenotype data into the study servers. On this example, the tables are on the demo Opal server as resources, which means they have to be loaded on the study server and be coerced as data frames; for more information read the [resourcer](https://cran.r-project.org/web/packages/resourcer/index.html) package documentation. Of course, the user may also have this information in Opal tables that can also be loaded into the servers using [standard DataSHIELD functions](https://isglobal-brge.github.io/resource_bookdown/datashield.html#demo). 

```{r}
library(dsBaseClient)
library(dsExposomeClient)
library(DSOpal)

builder <- newDSLoginBuilder()
builder$append(server = "server1", url = "https://opal-demo.obiba.org/",
               user = "administrator", password = "password",
               driver = "OpalDriver")
logindata <- builder$build()
conns <- datashield.login(logins = logindata)

datashield.assign.resource(conns, symbol = 'exposures', resource = list(server1 = 'EXPOSOME.exposures'))
datashield.assign.expr(conns, symbol = "exposures", expr = quote(as.resource.data.frame(exposures)))

datashield.assign.resource(conns, symbol = 'phenotypes', resource = list(server1 = 'EXPOSOME.phenotypes'))
datashield.assign.expr(conns, symbol = "phenotypes", expr = quote(as.resource.data.frame(phenotypes)))

ds.ls()
```

As it can be seen, there are two objects in the opal server  (`exposures` and `phenotypes`) that are the required ones for `dsExposomeClient` package. Note that both have a common variable ( `idnum` in our case) that will be used to link the two objects when creating the `ExposomeSet`

```{r}
ds.colnames("exposures")
ds.colnames("phenotypes")
```

Now, we are ready to create the  `ExposomeSet` object. To this end, the `ds.loadExposome ()` function should be used. It  takes as input parameters the names of the objects on the study server among some further configuration options. Please, refer to the function documentation for more information.

```{r}
ds.loadExposome(exposures = "exposures", 
                phenotypes = "phenotypes", 
                exposures.idcol = "idnum", 
                phenotypes.idcol = "idnum", 
                object_name = "exposome_object")

ds.class("exposome_object")
```


The function also allows to incorporate metadata on the exposures that can be used to improve results visualization. This requires to have another object (in our case called `description`) that can be loaded into the Opal server as

```{r load_description}
datashield.assign.resource(conns, symbol = 'description', resource = list(server1 = 'EXPOSOME.description'))
ds.class('description')
datashield.assign.expr(conns, symbol = "description", expr = quote(as.resource.data.frame(description)))
ds.class('description')
```

In our case this is the metadata we have

```{r}
ds.colnames("description")
```


Then, the `ExposomeSet` is created by adding further information to `dsloadExposome ()` function


```{r load_wiht_descr}
ds.loadExposome(exposures = "exposures",
                phenotypes = "phenotypes", 
                exposures.idcol = "idnum", 
                phenotypes.idcol = "idnum",
                description = "description",
                description.expCol = "Exposure",
                description.famCol = "Family",
                object_name = "exposome_object")

ds.class("exposome_object")
```

**NOTE:** In this vignette we are using this object that contains information on family exposures. All the following functions will also work if `description` is not provided since when `description = NULL` the `ExposomeSet` object will have the family name as the exposure one:

```
Exposure    Family
PFOA        PFOA
pm2.5       pm2.5
pm10        pm10
smoke       smoke
...               ....
```` 


